Decoupled Fallback: A Portable Single-Pass GPU Scan

### 2025-06-19 paper.tex
Add new diagrams, overhaul some old diagrams

\subsection{Evolution of Inter-Workgroup Scan Architectures}
Contemporary GPU scans depart from early PRAM-like scans~\cite{Horn, Hensley, Sengupta2006, Gress} by leveraging the GPU memory hierarchy, which consists of progressively faster but increasingly private tiers. Efficient scans integrate fine-grained intra-workgroup (\emph{local}) strategies in shared memory and registers with coarse-grained inter-workgroup (\emph{global}) strategies in global memory. Given their low arithmetic intensity, local scans are inherently memory-bound, as demonstrated by Merrill and Grimshaw~\cite{Merrill2009}. Consequently, recent scan optimizations have shifted toward refining global strategies, particularly inter-workgroup coordination and synchronization. This section reviews three key approaches---\emph{Scan-then-Propagate}, \emph{Reduce-then-Scan}, and \emph{Chained Scan}---analyzing their handling of inter-workgroup dependencies and efficiency in minimizing global memory access while maximizing bandwidth utilization.

@@

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\linewidth]{graphics/split.pdf}
  \caption{Splitting one 32-bit value into two 32-bit values with bit-packed tile state.}
\end{figure}

>>>>

\begin{figure*}[!t]
  \centering
  \makebox[\textwidth][c]{
    \begin{subfigure}[c]{0.48\textwidth}
      \centering
      \includegraphics[height=7.5cm,keepaspectratio]{graphics/ScanThenPropagate.pdf}
      \caption{\emph{Scan-then-Propagate.}}
      \label{fig:scan-then-propagate}
    \end{subfigure}
    \hfill
    \begin{subfigure}[c]{0.48\textwidth}
      \centering
      \includegraphics[height=6cm,keepaspectratio]{graphics/ReduceThenScan.pdf}
      \caption{\emph{Raking Reduce-then-Scan.}}
      \label{fig:reduce-then-scan}
    \end{subfigure}%
  }\\[1em]

  \makebox[\textwidth][c]{
    \begin{subfigure}[c]{0.48\textwidth}
      \centering
      \includegraphics[height=6cm,keepaspectratio]{graphics/StreamScan.pdf}
      \caption{\emph{Stream Scan.}}
      \label{fig:stream-scan}
    \end{subfigure}
    \hfill
    \begin{subfigure}[c]{0.48\textwidth}
      \centering
      \includegraphics[height=6cm,keepaspectratio]{graphics/DecoupledLookback.pdf}
      \caption{\emph{Decoupled Lookback.}}
      \label{fig:decoupled-lookback}
    \end{subfigure}%
  }

  \caption{The four global scan techniques described in \ref{sec:Evolution}. Dashed grey boxes indicate individual kernel launches.}
  \label{fig:four-techniques}
\end{figure*}

\subsection{Evolution of Inter-Workgroup Scan Architectures}
\label{sec:Evolution}
Contemporary GPU scans depart from early PRAM-like scans~\cite{Horn, Hensley, Sengupta2006, Gress} by leveraging the GPU memory hierarchy, which consists of progressively faster but increasingly private tiers. Efficient scans integrate fine-grained intra-workgroup (\emph{local}) strategies in shared memory and registers with coarse-grained inter-workgroup (\emph{global}) strategies in global memory. Given their low arithmetic intensity, local scans are inherently memory-bound, as demonstrated by Merrill and Grimshaw~\cite{Merrill2009}. Consequently, recent scan optimizations have shifted toward refining global strategies, particularly inter-workgroup coordination and synchronization. This section reviews three key approaches---\emph{Scan-then-Propagate}, \emph{Reduce-then-Scan}, and \emph{Chained Scan}---analyzing their handling of inter-workgroup dependencies and efficiency in minimizing global memory access while maximizing bandwidth utilization.

@@

\begin{figure*}[htbp]
  \centering
  \includegraphics[width=.95\textwidth]{graphics/DecoupledFallback.pdf}
  \caption{\emph{Decoupled Fallback}. For brevity, the chained propagation phase is condensed, see \ref{fig:decoupled-lookback} for the full behavior.}
  \label{fig:decoupled-fallback}
\end{figure*}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=1.03\linewidth]{graphics/split.pdf}
  \caption{Splitting one 32-bit value into two 32-bit values with bit-packed tile state.}
\end{figure}
